RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-07-03 background.tex
Section 1, 2.1, 2.2

In this section, we first introduce our running example, after which we introduce the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework. Lastly, we introduce argument schemes, and in particular, we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, which is an argument scheme that is used to form arguments and counter-arguments about situations involving goals. This will be our starting point in the next section.

@@

Most of the examples in this article, as well as the topic of discussion in the transcripts we analyze, come from the traffic simulator design exercise. In this exercise, designers are provided a problem description, requirements, and a description of the desired outcomes. The problem description is given in full in Appendix~\ref{ch:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering at UCI. It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory. To this end, a piece of software has to be developed in which students can create visual maps of an area, regulate traffic, and so forth. The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}. Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.

\subsection{Goal-oriented Requirements Language (GRL)}
\label{sect:background:grl}
GRL is a visual modeling language for specifying intentions, business goals, and \emph{non-functional requirements} of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}. GRL is part of the User Requirements Notation, an ITU-T standard, that combines goals and non-functional requirements with functional and operational requirements (i.e. use case maps) in one. GRL can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions. A GRL model is a connected graph of intentional elements that optionally are part of actors. All the elements and relationships used in GRL are shown in Figure~\ref{fig:grl_legend}.

@@

Figure~\ref{fig:trafficsim} illustrates a GRL diagram from the traffic simulator design exercise. An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system (\texttt{Student}, Figure~\ref{fig:trafficsim}), or the system itself ( \texttt{Traffic Tycoon}, Figure~\ref{fig:trafficsim}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g.  \texttt{Realistic simulation)} are often more related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are more related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}, Figure~\ref{fig:trafficsim}).

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{img/transcript_grl_0}
\caption{GRL Model for the traffic simulator.}
\label{fig:trafficsim}
\end{figure*}

Different links connect the elements in a GRL model. AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. In Figure~\ref{fig:trafficsim}, the goal  \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element. A contribution link has a qualitative contribution type or a quantitative contribution. Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the softgoal  \texttt{Dynamic simulation}. Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors. For example, actor  \texttt{Traffic Tycoon} depends on the actor  \texttt{Student} to perform the task  \texttt{Adjust car spawning rate} to fulfill its task  \texttt{Generate cars}.

GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$. Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution. GRL has a well-defined syntax and semantics, which are necessary if we want to incorporate it into a formal framework (requirements 1 and 2 as described in the introduction). Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details). GRL is also linked to Use Case Maps via URNLink ((\includegraphics[scale=1]{img/urnlink}) which provides traceability between concepts and instances of the goal model and behavioral design models. Multiple views and traceability are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.

GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements across multiple diagrams quantitatively, qualitatively or in a hybrid way. The satisfaction values from intentional elements in GRL can also be propagated to use case maps elements.  jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}. GRL also has the capability to be extended through metadata, links, and external OCL constraints. This allows GRL to be used in many domains without the need to change the whole modeling language. This feature also helps us to apply our argumentation to other domain such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.

The GRL model in Figure~\ref{fig:trafficsim} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model. However, the rationales and arguments behind certain intentional elements are not shown in the GRL model. Some of the questions that might be interesting to know about are the following:

\begin{itemize}
	\item Why does actor \texttt{Teacher} have only a single softgoal \texttt{Students learn from practice}? Why is this, for instance, not connected to any of the elements of \texttt{Student}?
	\item What does  \texttt{Adjust timing schemes of sensorless interactions} mean?
	\item Why does task \texttt{Keep same cars} contribut positively to \texttt{Realistic simulation} and negatively to \texttt{Dynamic simulation}?
	\item How does the \texttt{Student} control the \texttt{Traffic Tycoon}?
	\item Why does \texttt{Map design} have so many decompositions into other tasks?
\end{itemize}

These are the type of the questions that we cannot answer just by looking at the GRL models. The model in Figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships. In this article we aim to address this shortcoming.

@@

Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g. \cite{Raz1978-RAZPR,walton1990}) and artificial intelligence \cite{bratman1987,atkinson2007}. One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}. The idea is that an instantiation of such a scheme gives a presumptive argument in favor of, for example, taking an action. This argument can, then, be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.

A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the \emph{practical reasoning argument scheme} (PRAS). PRAS follows the following basic argument structure.

\begin{itemize}
\item[] We have goal $G$,
\item[] Doing action $A$ will realize goal $G$,
\item[] Which will promote the value $V$
\item[] \textit{Therefore} 
\item[] We should perform action $A$
\end{itemize}

So, for example, we can say that 
\begin{itemize}
\item[] We have goal  \texttt{Generate traffic},
\item[]  \texttt{Keep same cars} will realize goal  \texttt{Generate traffic},
\item[] Which will promote the value  \texttt{Simple design}
\item[] \textit{Therefore} 
\item[] We should perform action  \texttt{Keep same cars}
\end{itemize}

Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information. Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements. Some examples of critical questions are as follows.

\begin{enumerate}
\item Will the action bring about the desired goal?
%\item Does the goal promote the value stated?
\item Are there alternative ways of realizing the same goal?
\item Are there alternative ways of promoting the same value?
%\item Does doing the action have a side effect which demotes the value?
\item Does doing the action have a side effect which demotes some other value?
\item Does doing the action promote some other value?
%\item Does doing the action preclude some other action which would promote some other value?
\item Is the action possible?
\item Can the desired goal be realized?
\item Is the value indeed a legitimate value?
\end{enumerate}

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 4: if we find that  \texttt{Keep same cars} actually negatively influences the value  \texttt{Realistic simulation}, we have a counterargument to the above argument. Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3). For example, we can argue that  \texttt{Create new cars} also realizes the goal  \texttt{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}. In this section, we will briefly discuss the intuitions behind these semantics.}. Figure \ref{fig:pras:example} shows an argumentation framework with three arguments from the above example: the argument for  \texttt{Keep same cars} (A1), the argument for \texttt{Create new cars} (A3), and the argument that \texttt{Keep same cars} demotes the value  \texttt{Realistic simulation} (A2). The two alternative PRAS instantiations are A1 and A3. These arguments mutually attack each other, as \texttt{Keep same cars}  and \texttt{Create new cars} are considered to be mutually exclusive. Argument A2 attacks A1, as it points to a negative side-effect of \texttt{Keep same cars}.

\begin{figure}[ht!]
\centering
\begin{tikzpicture}
        \node[minimum size=1cm] (att3) [argNodeIN] at (0,-3) {$A_3$};
        \node[minimum size=1cm] (att1) [argNodeOUT] at (0,0) {$A_1$};
        \node[minimum size=1cm] (att2) [argNodeIN] at (3,0) {$A_2$};
         \path
    (att2) edge [attackLink] (att1)
    (att1) edge [attackLink,<->] (att3);
\end{tikzpicture}
\caption{Example argumentation framework.}
\label{fig:pras:example}
\end{figure}

Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics. The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated. In the argumentation framework in Figure~\ref{fig:pras:example}, argument A2 is undefeated because it has no attackers. This makes A1 defeated, because one of its attackers, A2, is undefeated. A3 is then also undefeated, since its only attacker, A1, is defeated by A2. Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$, corresponding to arguments for \texttt{Realistic simulation} and \texttt{Create new cars}.

\subsubsection*{Practical Argumentation and Goal Modeling}
\label{sect:background:pras:motivation}

Practical reasoning in the PRAS framework as described above provides a formal framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung1995} and its various extensions \cite{amgoud2002reasoning,modgil2009}. The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}. In this article, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model. This give a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it allows one to understand why certain other elements have been rejected.

@@

However, there are also some differences between PRAS and GRL. Not all elements and relationships of GRL fit into PRAS. For instance, PRAS does not have a notion of ``resource'', and many of the relationships of GRL do not occur in PRAS. Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson actually apply to GRL. Therefore, we develop our own set of argument schemes and critical questions in the next section by analyzing transcripts of discussions about the traffic simulator.

>>>>

In this section, we first introduce our running example, after which we introduce the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework. Next, we introduce argument schemes, and in particular, we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, which is an argument scheme that is used to form arguments and counter-arguments about situations involving goals. Finally, we give an overview of the integration between PRAS and GRL.  %This will be our starting point in the next section.

@@

We use the traffic simulator design case to explain the concepts and the framework in this paper. Traffic simulator design is the topic of discussion in the transcripts used for argumentation framework.   
%Most of the examples in this article, as well as the topic of discussion in the transcripts we analyze, come from the traffic simulator design exercise. 
In this exercise, designers are provided with a problem description, requirements, and a description of the desired outcomes. The problem description is given in full in Appendix~\ref{sect:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering courses at an American university. In order for the professor to teach students how various theories (such as queuing theory) around traffic lights work, a software analyst is hired to specify the goal and requirements of the system. %and It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory. 
To this end, a piece of software will be developed in which students can create visual maps of an area, regulate traffic, and so forth. The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning and it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}. Transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}. Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.

\subsection{Goal-oriented Requirements Language (GRL)}
\label{sect:background:grl}
GRL is a visual modeling language for specifying intentions, business goals, and \emph{non-functional requirements} of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}. GRL is part of the User Requirements Notation, an ITU-T standard, that combines goals and non-functional requirements with functional and operational requirements (i.e. use case maps) in one. GRL can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions. A GRL model is a connected graph of intentional elements that optionally are part of the actors. All the elements and relationships used in GRL are shown in Figure~\ref{fig:grl_legend}.

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{img/example_small}
\caption{Partial GRL Model of Traffic Simulator Example %(\todo{Marc}{Sepideh}{Turn into vector image, turn ``Dynamic simulation'' and ``Static Simulation'' into goals, turn ``Simulate'' into softgoal, add some resource}}
\label{fig:example-small}
\end{figure} %SG: I changed the figure a bit so that we won't need to change Figure 10. Let me know if you still want to change them according to the above.

@@

Figure~\ref{fig:example-small} illustrates a partial GRL diagram from the traffic simulator design exercise (the full figure is shown in Figure~\ref{fig:transcripts:grl} and will be discussed in Section~\ref{sect:gmas}). An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Tycoon}, Figure~\ref{fig:example-small}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g. \texttt{Show Simulation}) are often related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}).

Different links connect the elements in a GRL model. AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. In Figure~\ref{fig:example-small}, the goal \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element. A contribution link has a qualitative contribution type or a quantitative contribution. Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the task \texttt{Dynamic Simulation}. Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors. The goal \texttt{Generate Cars} depends on the goal \texttt{Value Has Changed}. %SG: I added it to the model but we can also remove it altogether. 
%\todo{Marc}{Sepideh}{Shall we add a dependency relation as well? They don't play a big role in our work..}

GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$. Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution. GRL has a well-defined syntax and semantics. Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details). GRL is also linked to Use Case Maps via URNLink ((\includegraphics[scale=1]{img/urnlink}) which provides traceability between concepts and instances of the goal model and behavioral design models. Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.

GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements across multiple diagrams quantitatively, qualitatively or in a hybrid way. The satisfaction values from intentional elements in GRL can also be propagated to the elements of Use Case Maps.  jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}. GRL also has the capability to be extended through metadata, links, and external OCL constraints. This allows GRL to be used in many domains without the need to change the whole modeling language. This feature also helps us apply our argumentation to other domains such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.

The GRL model in Figure~\ref{fig:example-small} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model. However, the rationales and arguments behind certain intentional elements are not shown in the GRL model. Some of the questions that might be interesting to know about are the following:

\begin{itemize}
	\item Why does actor \texttt{Traffic Tycoon} have only a single softgoal \texttt{Show Simulation}? Why is this, for instance, not connected to goal \texttt{Simulate}? %SG: I changed this question but let me know if you want to change them back and then change the graph.
	\item What does \texttt{Keep same cars} mean?
	\item Why does task \texttt{Keep same cars} contribute positively to \texttt{Static simulation} and negatively to \texttt{Dynamic simulation}?
	\item Why does \texttt{Generate Cars} XOR-decompose into two tasks?
\end{itemize}

These are the type of the questions that we cannot answer by just looking at GRL models. The model in Figure~\ref{fig:example-small} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships. In this article we aim to address this shortcoming.

@@

Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g. \cite{Raz1978-RAZPR,walton1990}) and artificial intelligence \cite{bratman1987,atkinson2007}. One approach is to capture practical reasoning using arguments schemes and critical questions~\cite{walton1990}. Applying an argument schemes results in an argument in favor of, for example, taking an action. This argument can, then, be tested with critical questions about, for instance, whether the action is possible given the situation. A negative answer to such a question leads to a counterargument to the original argument for the action.

Atkinson et al.~\cite{atkinson2007} develop an argument scheme for practical reasoning, which they call the \emph{practical reasoning argument scheme} (PRAS). This argument scheme is as follows\footnote{The original argument schemes uses slightly different terminology: we have replaced \emph{value} with \emph{softgoal}. Our results do not depend on these adjustments, but they make the relation between PRAS and GRL more clear.}:

\begin{itemize}
\item[] We have goal $G$,
\item[] Doing action $A$ realizes to goal $G$,
\item[] Which will contribute positively to the softgoal $S$
\item[] \textit{Therefore} 
\item[] We should perform action $A$
\end{itemize}

The capital letters $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals. Once instantiated, we obtain an argument. We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows:
\begin{itemize}
\item[] We have goal \texttt{Static simulation},
\item[] Doing action \texttt{Keep same cars} realizes goal \texttt{Static simulation}, 
\item[] Which contribute positively to the softgoal \texttt{Simulate} 
\item[] \textit{Therefore} 
\item[] We should perform action \texttt{Keep same cars}.
\end{itemize}

Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements. Some examples of critical questions are as follows.

\begin{enumerate}
\item Will the action realize the desired goal?
\item Are there alternative ways of realizing the same goal?
\item Are there alternative ways of contributing to the same softgoal?
\item Does performing the action contribute negatively to some other softgoal?
\item Is the action possible?
\item Can the desired goal be realized?
\item Is the softgoal indeed a legitimate softgoal?
\end{enumerate}

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 1: if we find that  \texttt{Keep same cars} actually does not realize goal \texttt{Static Simulation}, we can form a counter-argument.

Another way to elaborate an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that contributes to the same softgoal (question 3). For example,  can argue that another goal \texttt{Semi-dynamic Simulation} also realizes the softgoal \texttt{Simulate}. \todo{Marc}{Sepideh}{Do you think our example is good? I think it is confusing. Perhaps we should just create an artificial example here, loosely based on the traffic simulator but where the model makes sense to the reader. Now the tasks are all a bit vague and I think nobody knows what they actually mean.}

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}. Given an argumentation framework, once can compute very precisely which arguments are accepted and which are rejected.

\subsection{Combining PRAS and GRL}
\label{sect:background:pras:motivation}

The usefulness of argument schemes, and PRAS in particular, for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}. In this article, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model. This give a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it allows one to understand why certain other elements have been rejected.

@@

However, there are also some differences between PRAS and GRL. Not all elements and relationships of GRL fit into PRAS. For instance, PRAS does not have a notion of ``resource'', and many of the relationships of GRL do not occur in PRAS. Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson actually apply to GRL. Therefore, we develop our own set of argument schemes and critical questions in Section~\ref{sect:gmas} by analyzing transcripts of discussions about the traffic simulator. Before doing so, we give an overview of our framework in the next section.

### 2017-07-03 introduction.tex
Section 1, 2.1, 2.2

\subsection{Problem Statement}

A goal model is often the result of a discussion process between a group of stakeholders. For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background. Therefore, it is often not necessary to record all of the details of the discussion process that led to the final goal model. However, most real-world information systems -- e.g., air-traffic management, industrial production processes, or government and healthcare services -- are complex and are not constructed in a short amount of time, but rather over the course of several workshops. In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for various reasons.

First, it is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}. Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model. If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions. As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.

Secondly, other stakeholders, such as new developers on the team who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase. If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.

Thirdly, alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost. For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''. Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost. This leads to a poor understanding of the problem and solution domain. In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.

@@

The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with argumentation approaches techniques from Artificial Intelligence (AI) research\cite{atkinson2007}. We have identified five important requirements for our framework: 
\begin{enumerate}
\item The argumentation techniques should be close to actual discussions stakeholders or designers have in the early requirements engineering phase.
\item 
The framework must have formal traceability links between elements of the goal model and underlying arguments.
\item 
Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and visa versa.
\item 
The framework must have tool support.
\item 
A methodology should be identified that allows the framework to be used by practitioners.
\end{enumerate}

@@

The five requirements above are the success criteria of our approach, that is, the satisfaction of the six requirements will result in a positive answer to the research question.

The main contribution of this article is the RationalGRL framework,  which combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}. We develop a metamodel specifying how our framework extends GRL, and we use this as the basis for a prototype web-based implementation~\footnote{The tool is available at <GITHUB LINK>}. We develop an initial list of argument schemes and critical questions by analyzing a set of transcripts containing discussion about an information system. Our framework is fully extensible, meaning that argument schemes and critical questions can be added by users for specific problem domains. We develop a methodology for using RationalGRL, which consists developing goal models and posing arguments in an integrated way. This methodology is intended to be used by practitioners in the field. We formalize RationalGRL using propositional logic, and develop algorithms for adding goal elements and arguments.

\subsection{Organization}

This article is organized as follows. Section~\ref{sect:background} contains background and introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{}, and argument schemes. Section~\ref{sect:overview} provides a brief and high-level overview of our framework, together with a metamodel and the methodology. Section~\ref{sect:gmas} contains an in depth explanation of how we obtained an initial set of argument schemes by annotating transcripts from discussions about an information system, and in Section~\ref{sect:examples} we provide several examples of these arguments schemes. In Section~\ref{sect:formalframework} we formalize RationalGRL and our arguments schemes using propositional logic and we use well-known argumentation semantics to compute which arguments are accepted and which are rejected. We also develop various algorithms for the argument schemes in this section. In Section~\ref{sect:tool} we provide a brief overview of the prototype tool we developed for RationalGRL. Finally, Section~\ref{sect:discussion} contains a discussion, covering related work, future work, and a conclusion.

>>>>

\subsection{Problem Statement}
\label{sect:problem_statement}

A goal model is often the result of a discussion process between a group of stakeholders. For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background. Therefore, it is often not necessary to record all of the details of the discussion process that led to the final goal model. However, most real-world information systems -- e.g., air-traffic management, industrial production processes, or government and healthcare services -- are complex and are not constructed in a short amount of time, but rather over the course of several workshops. In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for various reasons stated below.

First, it is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}. Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model. If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions which results in having unnecessary discussions and revisions. As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.

Second, other stakeholders, such as new developers on the team who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase. If these users have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.

Third, alternative and different ideas and opposing views that could potentially have led to different goal diagrams could be lost. For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''. Thus, the resulting goal model merely contains the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost. This leads to a poor understanding of the problem and solution domain. In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.

@@

The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research\cite{atkinson2007}. We have identified five important requirements for our framework: 
\begin{enumerate}
\item The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
\item 
The framework must have formal traceability links between elements of the goal model and underlying arguments.
\item 
Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and vice versa.
\item 
The framework must have a tool support.
\item 
There should be a methodology for the framework to guide the practitioners in its application in real cases.
\end{enumerate}

The five requirements above serve as the success criteria of our approach. That is, if our framework satisfies these requirements, we reach our goal.

@@

%The five requirements above are the success criteria of our approach, that is, the satisfaction of the six requirements will result in a positive answer to the research question.

%SG: We need to write a bit about the framework in past and refer it here. 
To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL~\cite{}. The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}. %SG: we need to add 1-2 sentences about our previous work and then add what we do here... 
We develop a metamodel specifying how our framework extends GRL, and we use this as the basis for a prototype web-based implementation~\footnote{The tool is available at <GITHUB LINK>}. %SG: Placeholder for the link. 
We develop an initial list of argument schemes and critical questions by analyzing a set of transcripts containing discussion about an information system. Our framework is fully extensible, meaning that argument schemes and critical questions can be added by users for specific problem domains. We develop a methodology for using RationalGRL, which consists developing goal models and posing arguments in an integrated way. This methodology is intended to be used by practitioners in the field. We formalize RationalGRL using propositional logic, and develop algorithms for adding goal elements and arguments. We illustrates the steps of our framework  and how the framework can be applied in real cases with a traffic simulator example.

\subsection{Organization} %SG: Will get back to this at the end of the paper.

This article is organized as follows. Section~\ref{sect:background} contains background and introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, and argument schemes. Section~\ref{sect:overview} provides a brief and high-level overview of our framework, together with a metamodel and the methodology. Section~\ref{sect:gmas} contains an in depth explanation of how we obtained an initial set of argument schemes by annotating transcripts from discussions about an information system, and in Section~\ref{sect:examples} we provide several examples of these arguments schemes. In Section~\ref{sect:formalframework} we formalize RationalGRL and our arguments schemes using propositional logic and we use well-known argumentation semantics to compute which arguments are accepted and which are rejected. We also develop various algorithms for the argument schemes in this section. In Section~\ref{sect:tool} we provide a brief overview of the prototype tool we developed for RationalGRL. Finally, Section~\ref{sect:discussion} contains a discussion, covering related work, future work, and a conclusion.
